logistic regression

### 2021-07-13 scripts/43_comparisons_3.Rmd
update

@@

>>>>

simple <- read_csv('../data/shotgun_pheno/final_comprehensive_UPDATED_simple.csv')

@@

```{r}
# add a filtering step here 
# 50 and 25% 
pcts <- read_tsv('../data/shotgun_output/humann3/humann_final_out/pull_cr_d100_pcts_cpm_unstratified.tsv') %>% 
  gather('sampleid', 'cpm', names(.)[2]:names(.)[ncol(.)]) %>% 
  rename(pw = `# Pathway`)

keeppw <- pcts %>% 
  filter(cpm > 50) %>%
  ungroup() %>% 
  count(pw) %>% 
  filter(n > floor(nrow(simple) * 0.25)) %>% 
  pull(pw)

cts_fil <- pcts %>% 
  filter(pw %in% keeppw) %>% 
  spread('sampleid', 'cpm', fill = 0)

cts_fil %>% 
  write_tsv('../data/shotgun_output/humann3/humann_final_out/pull_cr_d100_pcts_cpm_unstratified_fil.tsv')
```

```{r}
# tox
pcts <- read_tsv('../data/shotgun_output/humann3/humann_final_out/pull_toxicity_pcts_cpm_unstratified.tsv') %>% 
  gather('sampleid', 'cpm', names(.)[2]:names(.)[ncol(.)]) %>% 
  rename(pw = `# Pathway`)

keeppw <- pcts %>% 
  filter(cpm > 50) %>%
  ungroup() %>% 
  count(pw) %>% 
  filter(n > floor(nrow(simple) * 0.25)) %>% 
  pull(pw)

cts_fil <- pcts %>% 
  filter(pw %in% keeppw) %>% 
  spread('sampleid', 'cpm', fill = 0)

cts_fil %>% 
  write_tsv('../data/shotgun_output/humann3/humann_final_out/pull_toxicity_pcts_cpm_unstratified_fil.tsv')
```

### 2021-07-13 scripts/48_16s_beta_diversity.Rmd
update

mp <- bc %>%
  as.data.frame() %>%
  rownames_to_column('sampleid') %>% 
  ungroup() %>% 
  inner_join(all_pheno) %>% 
  distinct(sampleid, .keep_all = T)  %>% 
  mutate(pool = str_extract(path_pool, 'Sample.+/'))

mp %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'pool', shape = 'center') +
  labs(title = 'PCOA of healthy and CART patients') +
  #theme_void() +
  ggsave('../figs/PCOA(bray-curtis) (ASV level)of CART patients_pool.pdf', width = 5, height = 6)
```

>>>>

mp <- bc %>%
  as.data.frame() %>%
  rownames_to_column('sampleid') %>% 
  ungroup() %>% 
  inner_join(all_pheno) %>% 
  distinct(sampleid, .keep_all = T)  %>% 
  mutate(pool = str_extract(path_pool, 'Sample.+/')) %>% 
  mutate(pool = str_replace(pool, 'Sample_','')) %>% 
  mutate(pool = if_else(str_detect(pool, 'IGO'), str_extract(pool, 'IGO.+$'), pool)) %>% 
  mutate(pool = str_replace(pool, '_1/|_comple.+$',''))

mp %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'pool') +
  labs(title = 'PCOA of CART patients') +
  #theme_void() +
  ggsave('../figs/PCOA(bray-curtis) (ASV level)of CART patients_pool.pdf', width = 5, height = 6)
```

### 2021-07-13 scripts/54_rethinking_binary.Rmd
back from my wonderful vacation
update

```{r}
library(rethinking)
library(tidyverse) 
library(ggpubr)
```

@@

# the std of b is 2, since we assume the microbiome could have big impact on immunology (based on our previous understanding and knowledge )
precis(mtox,  prob = 0.95)
plot(precis(mtox, depth = 2, prob = 0.95) ) 
# there is baseline difference between penn and Msk and MSK has higher tox rage in general.
# it doesn't seem diversity has a clear association with tox rate after correcting for center wise difference
div_value <- c(-1, 1)
```

@@

precis(mcr,  prob = 0.95)
plot(precis(mcr, depth = 2, prob = 0.95) ) 
# there isn't much difference in baseline level between the centers
# diversity seems to be positively associating with CR when the prior is too large sd
# if using appropriate prior the association is less sure

>>>>

```{r}
library(rethinking)
library(tidyverse) 
library(ggpubr)

@@

# the std of b is 2, since we assume the microbiome could have big impact on immunology (based on our previous understanding and knowledge )
precis(mtox,  prob = 0.95)
plot(precis(mtox, depth = 2, prob = 0.95 ) , pars = ('b'), labels = c('simpson_reciprocal'))
# there is baseline difference between penn and Msk and MSK has higher tox rage in general.
# it doesn't seem diversity has a clear association with tox rate after correcting for center wise difference
div_value <- c(-1, 1)

@@

precis(mcr,  prob = 0.95)
plot(precis(mcr, depth = 2, prob = 0.95 ) , pars = ('b'), labels = c('simpson_reciprocal'))
# there isn't much difference in baseline level between the centers
# diversity seems to be positively associating with CR when the prior is too large sd
# if using appropriate prior the association is less sure

### 2021-07-13 scripts/57_16s_lefse.Rmd
update

joined  %>%
  filter(grp == 'pull_cr_d100') %>% 
  ggboxplot(x = 'cr_d100', y = 'Relab', add = 'jitter', title = 'Outcome: cr_d100') +
  facet_wrap(direction ~ genus) +
  ggsave('../figs/amplicon/lefse_taxa_crd100.pdf', width = 15, height = 13)

@@

```{r}
joined  %>%
  filter(grp == 'pull_toxicity') %>% 
  ggboxplot(x = 'toxicity', y = 'Relab', add = 'jitter', title = 'Outcome: toxicity') +
  facet_wrap(direction ~ genus) +
  ggsave('../figs/amplicon/lefse_taxa_toxicity.pdf', width = 15, height = 13)
```

>>>>

joined  %>%
  filter(grp == 'pull_cr_d100') %>% 
  ggboxplot(x = 'cr_d100', y = 'Relab', add = 'jitter', title = 'Outcome: cr_d100') +
  facet_wrap(direction ~ genus, scales="free_y") +
  ggsave('../figs/amplicon/lefse_taxa_crd100.pdf', width = 15, height = 13)

@@

```{r}
joined  %>%
  filter(grp == 'pull_toxicity') %>% 
  ggboxplot(x = 'toxicity', y = 'Relab', add = 'jitter', title = 'Outcome: toxicity') +
  facet_wrap(direction ~ genus, scales="free_y") +
  ggsave('../figs/amplicon/lefse_taxa_toxicity.pdf', width = 15, height = 13)
```

### 2021-07-13 scripts/62_logistic_regression_version.Rmd
update

```{r}
library(magrittr)
library(ggeffects)
library(tidyverse)
library(lme4)
library(cowplot)
library(grid)
library(ggpubr)
```

@@

```{r}
m_tox <- glm(
  toxicity ~ simpson_reciprocal + center, 
  data = dat, 
  family = binomial(link = "logit")
)
summary(m_tox)

ggpredict(m_tox,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal')) %>% plot()
```

@@

```{r}
m_CR <- glm(
  cr_d100 ~ simpson_reciprocal + center, 
  data = dat, 
  family = binomial(link = "logit")
)
summary(m_CR)

ggpredict(m_CR,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal')) %>% plot()
```

@@

gtox <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    ) +
    ggsave(str_glue('../figs/amplicon/gtox_CLR_5.pdf'), width = 3, height = 8, dpi = 300)
```

@@

gCR <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    ) +
    ggsave(str_glue('../figs/amplicon/gcr_CLR_5.pdf'), width = 3, height = 8, dpi = 300)
```

>>>>

```{r}
library(magrittr) 
library(ggeffects)
library(tidyverse)
library(lme4)
library(cowplot)
library(grid)
library(ggpubr)
library(sjPlot)
theme_set(theme_sjplot())
```

@@

```{r}
m_tox <- glm(
  toxicity ~ simpson_reciprocal + center, 
  data = dat, 
  family = binomial(link = "logit")
)

summary(m_tox)

#ggpredict(m_tox,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal')) %>% plot()

# plot_model(m_tox, transform = NULL) +
#   ggsave('../figs/amplicon/diversity_tox_logi.pdf', height = 3, width = 4)

or_CI <- round(cbind(coef(m_tox), confint(m_tox)), digits=3) %>% 
  as.data.frame()

or_CI <- or_CI %>% 
  mutate(variable=rownames(or_CI)) %>% 
  rename(lower_bound = `2.5 %` , upper_bound = `97.5 %`)

# Now plot them
or_CI[2,] %>%  #remove row number 1 (The intercept)
  ggplot(aes(x = reorder(variable, V1), y = V1)) +
  geom_point(shape = 15,
             size  = 4, width = 0.1,
             position = "dodge", color="black") +
  geom_errorbar(aes(ymin  = lower_bound,
                    ymax  = upper_bound),
                width = 0.1,
                size  = 0.7,
                position = "dodge", color="turquoise4") +
  theme(axis.title = element_text(face = "bold")) +
  xlab("Variables") + ylab("Coefficient with 95% CI") +
  coord_flip(ylim = c(-0.5, 1)) +
  #geom_hline(yintercept = 1, color = "red", size = 1) +
   theme(axis.title = element_text(size = 0)) +
  theme(axis.text = element_text(size = 14)) +
  labs(title = 'Toxicity') +
  ggsave('../figs/amplicon/logis_div_tox.pdf', width = 5, height = 4)

```

@@

```{r}
m_CR <- glm(
  cr_d100 ~ simpson_reciprocal + center, 
  data = dat, 
  family = binomial(link = "logit")
)

or_CI <- round(cbind(coef(m_CR), confint(m_CR)), digits=3) %>% 
  as.data.frame()

or_CI <- or_CI %>% 
  mutate(variable=rownames(or_CI)) %>% 
  rename(lower_bound = `2.5 %` , upper_bound = `97.5 %`)

#Now plot them
or_CI[2,] %>%  #remove row number 1 (The intercept)
  ggplot(aes(x = reorder(variable, V1), y = V1)) +
  geom_point(shape = 15,
             size  = 4, width = 0.1,
             position = "dodge", color="black") +
  geom_errorbar(aes(ymin  = lower_bound,
                    ymax  = upper_bound),
                width = 0.1,
                size  = 0.7,
                position = "dodge", color="turquoise4") +
  theme(axis.title = element_text(face = "bold")) +
  xlab("Variables") + ylab("Coefficient with 95% CI") +
  coord_flip(ylim = c(-0.5, 1)) +
  #geom_hline(yintercept = 1, color = "red", size = 1) +
   theme(axis.title = element_text(size = 0)) +
  theme(axis.text = element_text(size = 14)) +
  labs(title = 'CR') +
  ggsave('../figs/amplicon/logis_div_cr.pdf', width = 5, height = 4)
```
```{r}
# plot_model(m_CR, transform = NULL, terms = c('simpson_reciprocal')) +
#   ggsave('../figs/amplicon/diversity_cr_logi.pdf', height = 3, width = 4)
# 
# get_model_data(m_CR, type = "est")
# log(get_model_data(m_CR, type = "est")$estimate)
```

```{r}
#ggpredict(m_CR,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal')) %>% plot()
```

@@

plot_model(g_tox)+
  ggsave('../figs/amplicon/genera_tox_logi_CLR.pdf', height = 3, width = 4)
```

```{r}
gtox_clr <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    )  
```

@@

plot_model(g_CR)+
  ggsave('../figs/amplicon/genera_cr_logi_CLR.pdf', height = 3, width = 4)
```

```{r}
gCR_clr <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    ) +
    ggsave(str_glue('../figs/amplicon/gcr_CLR_5.pdf'), width = 3, height = 8, dpi = 300)
```

# log 10 transformed version of the genera relab

```{r}
gmeta_log10 <- meta  %>% 
  inner_join(genera) %>% 
  select(toxicity,cr_d100, Akkermansia,Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus , center)

```

```{r}
g_tox <- glm(
  toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
  data = gmeta_log10, 
  family = binomial(link = "logit")
)
summary(g_tox)
gb <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
gr <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
ge <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
ga <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
gf <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
gtox_log <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    ) +
    ggsave(str_glue('../figs/amplicon/gtox_log10_5.pdf'), width = 3, height = 8, dpi = 300)

plot_model(g_tox) +
  ggsave('../figs/amplicon/genera_tox_logi_log10.pdf', height = 3, width = 4)
```

```{r}
g_CR <- glm(
  cr_d100 ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
  data = gmeta_log10, 
  family = binomial(link = "logit")
)
summary(g_CR)
gb <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
gr <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
ge <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
ga <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
gf <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
gCR_log <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    )

plot_model(g_CR)+
  ggsave('../figs/amplicon/genera_cr_logi_log10.pdf', height = 3, width = 4)
```

# side-by-side comparison of the CLR and log10 transformed

```{r}
tox_all <- ggarrange(gtox_log, gtox_clr, nrow = 1) +
    ggsave(str_glue('../figs/amplicon/gtox_all_5.pdf'), width = 6, height = 8, dpi = 300)

cr_all <- ggarrange(gCR_log, gCR_clr, nrow = 1) +
    ggsave(str_glue('../figs/amplicon/gCR_all_5.pdf'), width = 6, height = 8, dpi = 300)
```
